Reanalysis Step 2: Multiple Imputation (Clean Version)

### 2025-08-29 reanalysis-2-multiple_imputation_clean.qmd
refactor

# Variables to impute - binary
vars_impute_binary <- c(
  "cataract_surgery_bef_hitplus"
)

# Variables to impute - categorical
vars_impute_categorical <- c(
  "gender",
  "race_ethnicity_gp",
  "insurance_gp"
)

@@

# Comorbidity indicators (binary, recoded to 0/1)
vars_comorbidity <- c(
  "Glaucoma_bef_hitplus",
  "Otherretina_bef_hitplus"
)

@@

```{r prepare-data}
# Select variables and clean up
dr_cols <- c("PDR", "NPDR", "No_DR")
for (col in setdiff(dr_cols, names(df_original))) {
  df_original[[col]] <- NA_integer_
}
if (!"person_dr" %in% names(df_original)) {
  df_original[["person_dr"]] <- NA_integer_
}
df_mice <- df_original |>
  select(any_of(all_vars), any_of(dr_cols)) |>
  mutate(
    # Create person_dr from DR columns if needed
    person_dr = case_when(
      !is.na(PDR) & PDR == 1 ~ 2L,
      !is.na(NPDR) & NPDR == 1 ~ 1L,
      !is.na(No_DR) & No_DR == 1 ~ 0L,
      TRUE ~ person_dr
    ),
    # Convert to ordered factor
    person_dr = factor(person_dr, levels = c(0L, 1L, 2L), 
                      labels = c("No_DR", "NPDR", "PDR"),
                      ordered = TRUE),
    # Recode comorbidities to 0/1
    Glaucoma_bef_hitplus = ifelse(Glaucoma_bef_hitplus == "Present", 1, 0),
    Otherretina_bef_hitplus = ifelse(Otherretina_bef_hitplus == "Present", 1, 0),
    # Clean factor levels (remove spaces for formula compatibility)
    across(where(is.character), ~ factor(gsub("[^[:alnum:]_]", "_", .x)))
  ) |>
  dplyr::group_by(e_mrn_deidentified) |>
  dplyr::mutate(
    n_eyes = dplyr::n(),
    eye_index = dplyr::row_number(),
    fellow_baseline_VA = dplyr::if_else(n_eyes == 2, baseline_VA_logMAR[3 - eye_index], as.numeric(NA)),
    fellow_intermediate_VA = dplyr::if_else(n_eyes == 2, intermediate_VA_logMAR[3 - eye_index], as.numeric(NA)),
    fellow_outcome_VA = dplyr::if_else(n_eyes == 2, outcome_VA_logMAR[3 - eye_index], as.numeric(NA)),
    fellow_dr = dplyr::if_else(n_eyes == 2,
                               person_dr[3 - eye_index],
                               factor(NA, levels = levels(person_dr), ordered = TRUE))
  ) |>
  dplyr::ungroup() |>
  dplyr::select(-eye_index)

@@

# Variables that should always be included as predictors (key variables)
vars_always_predict <- c(
  "ever_lapse_binary",      # Exposure
  "baseline_VA_logMAR",     # Key outcome
  "outcome_VA_logMAR",      # Key outcome
  "age",                    # Demographics
  "gender",
  "CCI", "DCSI",           # Comorbidities
  "anti_VEGF"              # Key treatment
)

>>>>

# Variables to impute - binary
vars_impute_binary <- c(
  "catsurg_before_hitplus_cat"
)

# Variables to impute - categorical
vars_impute_categorical <- c(
  "gender_cat",
  "race_ethnic_cat",
  "insurance_cat"
)

@@

# Comorbidity indicators (binary factors from step 1)
vars_comorbidity <- c(
  "glaucoma_bef_hitplus_cat",
  "otherretina_bef_hitplus_cat"
)

@@

```{r prepare-data}
# Select variables and clean up (rely on step 1 outputs)
df_mice <- df_original |>
  select(any_of(all_vars)) |>
  mutate(
    # Ensure person_dr is ordered factor
    person_dr = if (is.numeric(person_dr)) {
      factor(person_dr, levels = c(0L, 1L, 2L), labels = c("No_DR", "NPDR", "PDR"), ordered = TRUE)
    } else {
      factor(person_dr, levels = c("No_DR", "NPDR", "PDR"), ordered = TRUE)
    },
    # Clean factor levels (remove spaces for formula compatibility)
    across(where(is.character), ~ factor(gsub("[^[:alnum:]_]", "_", .x)))
  ) |>
  dplyr::group_by(e_mrn_deidentified) |>
  dplyr::mutate(
    n_eyes = dplyr::n(),
    eye_index = dplyr::row_number(),
    fellow_baseline_VA = dplyr::if_else(n_eyes == 2, baseline_VA_logMAR[3 - eye_index], as.numeric(NA)),
    fellow_intermediate_VA = dplyr::if_else(n_eyes == 2, intermediate_VA_logMAR[3 - eye_index], as.numeric(NA)),
    fellow_outcome_VA = dplyr::if_else(n_eyes == 2, outcome_VA_logMAR[3 - eye_index], as.numeric(NA)),
    fellow_dr = dplyr::if_else(n_eyes == 2,
                               person_dr[3 - eye_index],
                               factor(NA, levels = levels(person_dr), ordered = TRUE))
  ) |>
  dplyr::ungroup() |>
  dplyr::select(-eye_index)

@@

# Variables that should always be included as predictors (key variables)
vars_always_predict <- c(
  "ever_lapse_binary",      # Exposure
  "baseline_VA_logMAR",     # Key outcome
  "outcome_VA_logMAR",      # Key outcome
  "age",                    # Demographics
  "gender_cat",
  "CCI", "DCSI",           # Comorbidities
  "anti_VEGF"              # Key treatment
)
